---
title: "SIBB512: Assignment"
author: "Bhoom Suktitipat"
date: "Nov/28/2019"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Instruction

1. Read through this assignment and fill in the code chunks for the answers.
2. Modify the author name and `knit` the file to create your answer.
3. Save the result knitted file using the following format "SIBB512-FirstName-Lastname.html"
4. Send it to bhoom.suk@mahidol.edu with the subject line "SIBB512 R Assignment"

## Import Data

Border crossing entry data from [Kaggle](https://www.kaggle.com/akhilv11/border-crossing-entry-dataBorder_Crossing_Entry_Data) has been saved to the current workspace. You can find the file in the File Panel on the lower right part of the screen. 

To import the dataset, you can use the GUI menu at the upper right panel, *click* "Import Dataset". I preferably recommend using _"From text (readr)"_ as the package came with several automate import help.

Please follow the import dialog box, change the type of delimitter if the data preview does not look correct. 

Use the "History" Panel to copy the code used to import data into this file.

```{r border_data}
require(readr)
Border_Crossing_Entry_Data <- readr::read_csv("Border_Crossing_Entry_Data.csv")
```

After importing the data, you should be able to see the *Data* named `Border_Crossing_Entry_Data` in the Environment panel (upper right corner).

This is a table similar to a spread sheet in Excel. In R, this is a common format convenient for data analysis called "Data Frame".

Can you change the name of the data to `border_data`?
[Hint: Try to re-import the data, but change the `Name` in the import option dialog box to `border_data`]

## Getting to know your data

How may columns are there? [Try click on the data to preview it.]
[Hint: Insert code chuck and try ncol() and names() function] What's the difference between the two function?
```{r ncol}
border_data <- Border_Crossing_Entry_Data
ncol(border_data)
```

```{r names}
names(border_data)
```

How many observations are there?
[Hint: nrow() and dim()]
```{r dimension}
border_data %>%
  nrow()
```

```{r dim_dimension}
dim(border_data)
```

Can you guess what type of info the column name "Value" is telling us?

How many mode of border crossing was recorded?

## Subsetting data
What was the most common method of border crossing from California?
[Hint: Try `table(border_data$Measure)`]
```{r ans01}

```
Can you get the same answer using `%>%` and `count()` in `tidyverse`?
```{r ans02}

```

What's the benefit of piping? 
- can you time the operation?
```{r ans03}

```

### which?

How many records are related to "California"?
[Hint 1: ?subset]
[Hint 2: ?which]


```{r californian}
californian <- subset(border_data, State == "California")
# californian <- border_data[which(border_data$State=="California"), ]
```

Can you count the numbers of records related to each State within the data?
[Hint: use `group_by()` and `count()` in tidyverse]
```{r ans04}

```

### which & and
How many people walk acorss the border from California in this data?
```{r californian_walker}
californian_walker <- californian[which(californian$Measure=="Pedestrians"),]
```
You can get the summation of the *Value* using `sum(californian_walker$Value)` or use Pipe operation. Pipe calculate the same answer without the need to save an intermediate data frame in the memory.
```{r ans05}

```

## Count by category

Try help function `?tapply` 
```{r crosser_count}
californian_crossers <- tapply(californian$Value, californian$Measure, sum)
californian_crossers
```
Can you use `group_by()` to get similar answer?
```{r ans06}
californian %>% 
  group_by(Measure) %>% 
  summarise(Number=sum(Value)) %>% 
  arrange(desc(Number))
```

What is the most common *Port* for Californian to cross the border?
```{r ans06}

```

Can you create a plot to represent the mode and number of border crossing for Californian?
```{r ans07}

```
