---
title: "Exploratory Data Analysis"
output: 
  html_notebook: 
    toc: yes
---

# Overview

The content of this markdown notebook was adopted from [R for Data Science](https://r4ds.had.co.nz/index.html) book by Hadley Wickham. The book is available online for free. 

We will learn how to to filter, arrange, summarise, mutate and visualize your data with dplyr and ggplot2 using the Titanic data!

```{r}
# install.packages("tidyverse")
library(tidyverse)
```

# EDA in Brief

**GOAL** Try to find variation and covariation within the dataset. 

What measrement/statistics can you use to measure variation and covariation among the variables within the dataset?

## The *Diamond* dataset.
To get information about a specific dataset included with R packages. For example, the *diamonds* dataset included with **ggplot2** package. Try the following command.
```{r}
?diamonds
```
### What information is given by the variable `cut`?

### Visualizing Variation

#### How do you visualize the variation in the diamond data as described by _cut_?
```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

### How do you interpret the graph above?

### Can you give the exact count by cut and display the result in a table?
```{r}
diamonds %>% count(cut)
```

#### Can you show the variation by carat?
```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat))
```
### histogram binwidth
How does the binwidth affect the shape of the data? Try vary the binwidth to "1" and "0.1" (Hint: `?geom_histogram`)
What binwidth gives us more detail on the variation in the diamond size?

## Typical values?

Which diamond sizes are the most common? Can you guess why?

## Unusual values?

Is there an unusually small diamond within the diamond data set?

### How do you visualize these outliers?
Name 2 type of graphs that can help us achieve this goal.
1. 
2. 

How do you create the above two plots?

# Titanic Data

Get the data from [here](https://www.kaggle.com/c/titanic)
- `train.csv`: Data for model training
- `test.csv`: Data for model testing
- `gender_submission.csv`: result submission for Kaggle competition

## Import Train Data
```{r}
passengers <- read.csv("train.csv")
```

View the data by either clicking on the Data name on the panel. or Use the command `View(passengers)` or `head(passengers)`. 
Let's look at the data!

```{r}
head(passengers)
```

## General summary of the data
```{r}
summary(passengers)
```

With the `tidyverse` package you can pipe your data to a command. For example, to summarize the `passengers` data. You can pipe it with `%>%` to the function `summary()`.
```{r tidy_summary}
passengers %>% 
  summary()
```

You can combine pipe. 
For example, removing any records with `NA` first, then summarize the data. 
```{r}
passengers %>% 
  drop_na() %>%
  summary()
```
You can see that the male:female changed from 577:314 to 453:261. 

## Filter

Choose specfic rows based on specific criteria. 
```{r}
passengers %>%
  filter(Sex == "male") 
```

## Arrange the data

Try to order the data by Fare. In this example, to put the most expensive at the top, we can use `desc()`. 
```{r}
passengers %>% 
  arrange(desc(Fare))
```

## Create new variable

`mutate()` the old dataset to include a new variable `famsize`. You know that the variable `Parch` is the number of parents and children while `SibSp` is the number of siblings and spouses. You can add these together to get a new variable `FamSize`.
```{r}
passengers %>%
  mutate(FamSize = Parch + SibSp)
```

## Convert variable from number to text
Turn numerical values of Survived column to "No" & "Yes" (new data frame)
```{r}
passengers1 <- passengers %>%
  mutate(Survived = ifelse(Survived == 0, "No", "Yes"))
```

## Plotting your data

### bar plot
```{r}
ggplot(passengers , aes(x = Sex)) + 
  geom_bar()
```

What's your interpretation of the graph?

### Scatter Plot

Try looking for relationship between `Fare` and `Age`. We will try to make a point first.
```{r}
ggplot(passengers %>% drop_na(), aes( x = Age, y = Fare)) +
  geom_point()
```

### Adding color to the plot

We can add additional insight to the graph by providing color. 
```{r}
ggplot(passengers %>% drop_na(), aes( x = Age, y = Fare, color = Sex)) +
  geom_point()
```

### Adding Facet to the Data

What about `Fare`, `Age` and `Sex` impact on `Survival`?
```{r}
ggplot(passengers1 %>% drop_na(), aes(x = Age, y = Fare, color = Sex)) +
  geom_point() +
  facet_grid(~Survived)
```

### More meaning bar plot
Giving info on sex and survival. But with bar plot, we use `fill` option to fill in color by categorical variable instead of using `color` to fill each data point.
```{r}
ggplot(passengers1, aes(x = Sex, fill = Survived)) +
  geom_bar()
```

## Summarizing and Grouping Data

### summarise
Finally, to calculate some aggregate statistics from the data, we can use `summarise()`.
```{r}
passengers %>%
  summarise(MeanFare = mean(Fare))
```

Can you calculate the median fare?

What about the meadian fare that male passengers had to pay? 
```{r}
passengers %>%
  filter(Sex == "male") %>%
  summarise(MeanFare_male = mean(Fare))
```
What about the meadian fare that female passengers had to pay? 

How to find out the total number of female that survived together with the `MeanFare`?
```{r}
passengers %>% 
  filter(Sex == "female") %>%
  summarise(MeanFare = mean(Fare), NumSuv = sum(Survived))
```

### group_by
Instead of calculating the median fare twice for male and female, we can use `group_by()`
```{r}
passengers %>% 
  group_by(Sex) %>%
  summarise(MedianFare = median(Fare))
```

### sum
```{r}
passengers %>% 
  group_by(Sex) %>%
  summarise(MedianFare = median(Fare), NumSurv = sum(Survived))
```